The function--sheaf dictionary for smooth commutative group schemes over finite fields

### 2014-04-03 dictionary.tex
Added the Hochschild-Serre spectral sequence

\begin{lemma}\label{prop:etale-2cocycle}
    If $G$ is an \'etale commutative group scheme over $\Fq$
    then the homomorphism
    \[
    \QCiso{G} \to \Hh^0(\Weil{}, \Hh^2(G(\bFq),\EEx)),
    \]
    defined in Section~\ref{sec:invisible}, is surjective.
\end{lemma}

\begin{proof}
Pick $[g]\in \Hh^0(\Weil{}, \Hh^2(G(\bFq),\EEx))$. Then $g\in Z^2(G(\bFq),\EEx)$ is fixed by Frobenius, so $\,^{\Frob{}}g = g h$ for some $h\in B^2(G(\bFq),\EEx)$. Accordingly, $h(x,y) = \frac{f(x+y)}{f(x) f(y)}$ for some $f : G(\bFq) \to \EEx$, so
\begin{equation}\label{gf}
g(\Frob{}(x),\Frob{}(y)) = g(x,y) \frac{f(x+y)}{f(x) f(y)}
\end{equation}
for all $x,y\in G(\bFq)$.
Set $\gqcs{L}_x = \EE$ for every $x\in G(\bFq)$ and define $\mu_{x,y} : \gqcs{L}_{x+y} \to \gqcs{L}_x \otimes \gqcs{L}_y$ by $\mu_{x,y}(1) = g(x,y) 1 \otimes 1$.
Since $g\in Z^2(G(\bFq),\EEx)$, the indexed set of isomorphisms $\{ \mu_{x,y} \tq x,y \in G(\bFq) \}$ satisfies \ref{qc.2}.
Define $\phi_x : \gqcs{L}_{\Frob{}(x)}\to \gqcs{L}_x$ by $\phi_x(1) = f(x)$.
Then set of isomorphisms $\{ \phi_{x} \tq x \in G(\bFq) \}$ satisfies \ref{qc.3} by \eqref{gf}. In this way, the cocycle $g\in Z^0(\Weil{}, Z^2(G(\bFq),\EEx))$ defines a quasicharacter sheaf $\qcs{L}$. It is now clear that $\qcs{L}$ maps to $[g] \in \Hh^0(\Weil{}, \Hh^2(G(\bFq),\EEx))$ under \eqref{2cocycle}.
\end{proof}

@@

\begin{lemma}\label{prop:etale-trace}
If $G$ is an \'etale commutative group scheme over $\Fq$
then for every the $\chi \in G(\Fq)^*$ there is a visible quasicharacter sheaf $\qcs{L}$ on $G$ such that $\trFrob{\qcs{L}} = \chi$.
In particular, Trace of Frobenius
\[
\TrFrob{G} : \QCiso{G} \to G(\Fq)^*,
\]
is surjective, even when restricted to visible quasicharacter sheaves.
\end{lemma}

@@

\begin{proof}
\todo{Clifton, you are here.}
\end{proof}

>>>>

\begin{lemma}\label{lem:etale-2cocycle}
    If $G$ is an \'etale commutative group scheme over $\Fq$
    then the group homomorphism
    \[
    \QCiso{G} \to \Hh^0(\Weil{}, \Hh^2(G(\bFq),\EEx)),
    \]
    defined in Section~\ref{sec:invisible}, 
    is surjective with a canonical section.
\end{lemma}

\begin{proof}
Pick $[g]\in \Hh^0(\Weil{}, \Hh^2(G(\bFq),\EEx))$. Then $g\in Z^2(G(\bFq),\EEx)$ is fixed by Frobenius, so $\,^{\Frob{}}g = g h$ for some $h\in B^2(G(\bFq),\EEx)$. Accordingly, $h(x,y) = \frac{f(x+y)}{f(x) f(y)}$ for some $f : G(\bFq) \to \EEx$, so
\begin{equation}\label{gf}
g(\Frob{}(x),\Frob{}(y)) = g(x,y) \frac{f(x+y)}{f(x) f(y)}
\end{equation}
for all $x,y\in G(\bFq)$.
Set $\gqcs{L}_x = \EE$ for every $x\in G(\bFq)$ and define $\mu_{x,y} : \gqcs{L}_{x+y} \to \gqcs{L}_x \otimes \gqcs{L}_y$ by $\mu_{x,y}(1) = g(x,y) 1 \otimes 1$.
Since $g\in Z^2(G(\bFq),\EEx)$, the indexed set of isomorphisms $\{ \mu_{x,y} \tq x,y \in G(\bFq) \}$ satisfies \ref{qc.2}.
Define $\phi_x : \gqcs{L}_{\Frob{}(x)}\to \gqcs{L}_x$ by $\phi_x(1) = f(x)$.
Then set of isomorphisms $\{ \phi_{x} \tq x \in G(\bFq) \}$ satisfies \ref{qc.3} by \eqref{gf}. In this way, the cocycle $g\in Z^0(\Weil{}, Z^2(G(\bFq),\EEx))$ defines a quasicharacter sheaf $\qcs{L}$. It is clear that $\qcs{L}$ maps to $[g] \in \Hh^0(\Weil{}, \Hh^2(G(\bFq),\EEx))$ under \eqref{2cocycle}.\todo{Canonical section?} 
\end{proof}

@@

\begin{lemma}\label{lem:etale-trace}
If $G$ is an \'etale commutative group scheme over $\Fq$
then for every the $\chi \in G(\Fq)^*$ there is a visible quasicharacter sheaf $\qcs{L}$ on $G$ such that $\trFrob{\qcs{L}} = \chi$.
In particular, Trace of Frobenius
\[
\TrFrob{G} : \QCiso{G} \to G(\Fq)^*,
\]
is surjective, even when restricted to visible quasicharacter sheaves.
\end{lemma}

@@

\begin{proof}
The Hochschild-Serre spectral sequence (for finitely generated abelian groups!) provides decomposition 
\[
\Hh^2(G(\bFq),\EEx)  \iso \Hh^1(\Weil{}, \Hh^1(G(\bFq)\rtimes \Weil{},\EEx)) \oplus \Hh^0(\Weil{}, \Hh^2(G(\bFq),\EEx))
\]
and Lemmas~\ref{lem:etale-2cocycle} and \ref{lem:etale-trace} provide the maps from $\QCiso{G}$ pictured below:
\[
\begin{tikzcd}
\ &&  \QCiso{G} \arrow[swap]{dl}{\TrFrob{G}} \arrow{dr}{ S_{G} } && \\
0 \arrow{r}{} &  \Hh^1(\Weil{}, \Hh^1(G(\bFq),\EEx)) \arrow{r}{} & \Hh^2(G(\bFq)\rtimes\Weil{},\EEx) \arrow{r}{} & \Hh^0(\Weil{}, \Hh^2(G(\bFq),\EEx)) \arrow{r}{} & 0.
\end{tikzcd}
\]

\todo{Clifton, you are here.}
\end{proof}
